Periodicity of Geodesics on the Necker Cube Surface

### 2017-12-04 newpaper.tex
120417

\title{Periodicity of Geodesic Flows on the Necker Cube Surface}
\author{Pavel Javornik}

@@

\begin{abstract}
We study dynamical properties of geodesic flows on a flat, periodically constructed Euclidean cone surface by obtaining an infinite-type translation cover branched over conical singularities, of which admits a flat induced metric and a connection between any two points in the tangent vector bundle of the surface has trivial holonomy. A unit vector flow on the unit tangent bundle of such a surface is well-defined with a canonical vector representation in $\RR^2$. We study this infinite-type surface as a $\ZZ^2$ cover of a compact Veech surface belonging to the stratum of $\cH(2,2,2,2)$ translation surfaces, and use its $SL(2,\RR)-$commensurable Veech group to prove results that relate directions of flows to the periodicity and ergodicity of their lifts in the cover.
\end{abstract}
\begin{figure}[H]
\centering
\includesvg[width=2.2in]{cubesxyzpaths2}
\label{fig:front}
\caption{Periodic and drift-periodic flows on the Necker cube surface}
\end{figure}

@@

Historically, the Necker cube[citation] has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems prior to it being popularized in the works of illusionary artist M.C. Escher (pictured below). The crystallographer Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometry[cite]. The solid presentation of the cube, when rendered as a flat surface with its back faces removed, achieves a similar effect when its three visible faces are shaded in a particular way. We are interested in the infinite tiling of this structure (figure $\ref{fig:front}$), which has the appearance of a rhombile tiling of the Euclidean plane, and refer to it as the Necker cube surface.  Video game fans might even recognize it as the same surface that appears in the 1982 arcade game by Gottlieb, ``Q*bert."

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{escher.jpg}
\includegraphics[scale=0.55]{escher2.jpg}
\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and impressed into a banner in ``Convex and Concave."[citation]}
\label{fig:Escher}
\end{center}
\end{figure}

Flat, periodic surfaces such as the Necker cube surface are of much interest to differential geometers for two reasons in particular. The first is that, locally, neighborhoods are isometric to the plane, making it remarkably easy to describe parallel transports between two points in tangent vector bundles over the surface by linear group actions. The second is that Riemann surfaces constructed out of infinitely many polyhedra can be \emph{flattened} and contained in the plane in such a way that the metric is preserved. The advantage in this case is that one obtains something akin to infinite billiard tables where a geodesic flows are represented by series of straight line segments in the plane. This paper will use such methods to prove dynamical results about Geodesic flows on the Necker cube surface.

@@

\subsection{Discussion of Results}
Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object. Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane. Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$. $\ref{fig:front}$.

\begin{wrapfigure}{l}{2.2in.}
\includesvg{vectorplane}
\end{wrapfigure}
\noindent Let $x_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at a point $u_0$. Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0. Call this projected vector $v_0\in\RR^2$.
\\\\
The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface which we denote by $\Phi_t:\RR\rightarrow\bUs$. A rational initial trajectory in $\RR^2$ falls into one of two categories:
\begin{Def}
Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$. We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd. We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$. We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity. We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
\end{Def}
\noindent This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface. Denote $\bUs$'s translational symmetries by $\text{Trans}(\bUs)$. Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T)=p(\Phi(t))$ for all $t\in\RR$. In either case we call $T$ the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies these criteria. We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, for which geodesic behavior is non-deterministic.

\begin{thm*}
(Dynamical Properties of a Geodesic on $\bUs$) Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v v_0\in\RR^2$. Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item $\Phi$ is periodic with period $T\in\RR$ if  $v_0\in\cO$.
\item $\Phi$ is drift-periodic with period $T\in\RR$ if $v_0\in\cE$.
\end{enumerate}
\end{thm*}

The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
\\
\compav{theorem about most directions being recurrent?}

@@

\section{Periodic Tiling of Necker Cubes}
This section will detail how the Necker Cube surface is obtained as an infinite regular cover of a branched torus. Consider the Necker cube with the following identifications:
\begin{figure}[H]
\centering
\includesvg[width=1.6in.]{neckercube}\includesvg[width=2in.]{neckercubepath}
\end{figure}

We denote this surface $\bG$. $\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a unit cube. $\bG$ is a genus one surface with three vertices homeomorphic to $\TT^2$. Check that every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$. These are the conical singularities of the surface, $\Sigma$. We denote the surface without its singularities $\bG\backslash\Sigma = \bGs$. Thus, $\pi_1(\bGs)\cong\pi_1((\TT^2 \text{ with 3 punctures}))\cong\pi_1(S^1\vee S^1\vee S^1\vee S^1)\cong\FF_4$, the free group of four generators. We may flatten $\bG$ and $\bGs$ into an L-shaped torus: \\
\begin{figure}[H]
\centering
\includesvg[width=3in]{LTorus}
\end{figure}
\noindent This L-shaped torus is then cut along the dotted lines and pieced together to resemble a $2\times2$ torus with a unit square removed from the center:\\
\begin{figure}[H]
\centering
\includesvg[width=2in]{BranchTorus}
\caption{$\bGs$ with 4 independent, closed paths labeled identified}
\end{figure}
Denote its fundamental group by $\pi_1(\bGs)=\<a,b,c,d\>$, where $a,b,c,d$ are the labeled independent paths on the surface. Consider the following map:

\begin{Def}
Let $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ be the group homomorphism defined on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$
\end{Def}

\begin{Def}
The punctured \textbf{Necker Cube Surface} is a regular cover of $\bGs$, and is denoted $\bUs$. It's fundamental group presentation is the subgroup of $\pi_1(\bGs)$, $\pi_1(\bUs)=\ker\varphi_1$.
\end{Def}

Thus $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong \varphi_1(\pi_1(\bGs))=\ZZ^2$ is the associated \emph{deck group} of the cover. $\bUs$ is the primary object of study.

\subsection{Deck Transformations of the $\ZZ^2$ Cover}
The set of Deck transformations of the covering map $\bUs\twoheadrightarrow\bGs$ is isomorphic to the deck group $\ZZ^2$ that we induce from $\text{Trans}(\CC)$. First we remove open unit squares on an even integer lattice in the plane:\\

\noindent  Let $m,n\in\ZZ$.
\begin{equation}
\label{eq:P}
\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:~|u-2m|<\frac{1}{2},~|v-2n|<\frac{1}{2},~u=2m\pm\frac{1}{2}~v=2n\pm\frac{1}{2} \big{\}}.
\end{equation}

\noindent Now make the following identifications on $\mathbf{P}:$
\\ Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be families of paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (\frac{-1}{2},\frac{1}{2})$. $\bUs$ is recovered as a topological quotient by making the following edge identifications:

\begin{equation}
\begin{split}
(\zeta(t))\sim(e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph})\\
(\eta(t))\sim_(e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph})
\label{eq:rel2}
\end{split}
\end{equation}

These spaces and identifications could be adapted to recover $\bGs$ by restricting $\mathbf{P}$ to a $2\times2$ square about the origin, only allowing $m=n=0$, and identifying the outer edges of the square much like a 2-torus. The translations $+2,+2i\in \text{Trans}(\CC)$ induce horizontal and vertical maps $f_h,f_v\in\text{Trans}(\bUs)$. The group $\<f_h,f_v\>$ is isomorphic to $\ZZ^2$.

@@

\subsection{Monodromy Group Representation}
In this section we describe a translation cover of $\bGs$ which and $\bUs$ which is factored through as the kernel of a homomorphism from $\pi_1(\bGs)\rightarrow \SO(2,\ZZ)$. This homomorphism represents the monodromy group as the fundamental group acting on the unit tangent bundles $T^1\bGs$ and $T^1\bUs$.

\begin{Def}
Let $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ be the group homomorphism defined on the generators $a,b,c,d$:
\begin{align*}
a,b & \mapsto \begin{pmatrix}
1 && 0 \\ 0 && 1\end{pmatrix}=I_2\\
c & \mapsto\begin{pmatrix}0 && -1\\1 && 0\end{pmatrix}=R\\ d & \mapsto\begin{pmatrix}0 && 1\\-1 && 0\end{pmatrix}=R^3
\end{align*}
\end{Def}

Recall that a \emph{translation surface} is a compact, locally flat surface that admits a discrete number of conical singularities and has trivial linear holonomy on any arbitrary path. If the total cone angle of the singularity is some integer multiple of $2\pi$, $2\pi(d+1)$ for $d\geq 0$, we say that singularity has degree $d$. Two translation surfaces belong to the same stratum, or family, of surfaces characterized by the number and degree of conical singularities. Consider the effect that paths $c,d$ have on holonomy:

\begin{figure}[H]
\centering
\includegraphics[width=2.5in]{monodromy.png}
\caption{Effect a nontrivial loop has on arbitrary basis vectors (i,j).}
\label{fig:loop}
\end{figure}

We construct a cover of $\bGs$ by taking four copies of it and cyclically pairing edges, and show that this cover has trivial holonomy. Denote the unit tangent bundle of $\bGs$ by $T^1\bGs$. Because $\bGs$ can be contained in the plane, we consider vectors tangent to $\bGs$ as vectors in $\RR^2$. Let $x_0\in\bGs$. The fiber over $x_0$ under the projection $T^1\bGs\twoheadrightarrow\bGs$ is the set $E_{x_0}$. Let $\triangledown$ be a flat metric connection on $T^1\bGs$ and $p=(x_0,v_0)\in E_{x_0}$ be a single element in the fiber over $x_0$ with $v_0\in\RR^2$. We denote the \emph{holonomy group} of the connection with base point $p$ by $Hol_p(\triangledown)=\{A\in SO(E_{x_0})~:~ p = A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$. An element in $SO(E_{x_0})$ acts on $p$ by rotation of the vector $v_0$. Denote the \emph{holonomy bundle} based at $p$ by $H(p)=\{p'\in E_{x_0}~:~ p'= A\cdot p \}$, the set of elements in the orbit of $p$ where a closed loop $\gamma:[0,1]\rightarrow\bGs$ defines a horizontal lift to $\gamma^*:[0,1]\rightarrow T^1\bGs$ by parallel transport of $v_0$ along the path such that $p=\gamma^*(0)$, and $p'=\gamma^*(1)$. Since $\bGs$ is flat and trivial paths have trivial holonomy, $Hol_p(\triangledown)$ acts on $H(p)$ monodromy group is discrete and well-defined on $\bGs$.

\begin{lem}
$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is a monodromy representation of the surface.
\begin{proof}
It is clear from Figure 6 that this is the case as concatenating a closed loop with paths $c$ or $d$ would rotate a vector exactly by the cone angles of the singularities that they loops around: $\pm \frac{3\pi}{2}$.
\end{proof}
\end{lem}

@@

\subsection{A Four-Fold Cover}

\begin{Def}
Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
\end{Def}

\compav{Prove that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$}

$\btUs$ is also a cover of $\bMs$ with fundamental group $\ker\varphi_1\big\vert_{\pi_1(\bMs)}$. In the former case the deck group of the cover is $\SO(2,\ZZ)$, while in the latter case it is $\ZZ^2$. The deck group of the cover $\btUs\rightarrow\bGs$ is $\ZZ^2\rtimes\SO(2,\ZZ)$. \compav{worth talking about?}. $\btUs$ is a four-fold cover of $\bUs$ with trivial holonomy visualized as four copies of $\bUs$, which can be thought of as countably many copies of . We can identify a closed path on $\bGs$ as an element of the deck group $\ZZ^2\times \SO(2,\ZZ)$ or, equivalently, $ 2\ZZ+2\ZZ i\times \ZZ/4\ZZ$ if we want to consider a metric on the ambient space $\CC$ that is consistent with the tiling of $\bGs$. Further, we can rotate everything in $\CC\times \ZZ/4\ZZ$ along with $\btUs$ according to the element of $\SO(2,\ZZ)$ that acts non-trivially on the holonomy of $\bUs$. For example, let $P\in2\ZZ+2\ZZ i$ and consider the following curve on $\bUs$:

@@

\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group isomorphic to $2\ZZ+2\ZZ i \times \ZZ/4\ZZ$. Observe that it will only close if the curve acts trivially on the holonomy of $\bUs$:

\begin{figure}[H]
\centering
\includegraphics[width=3in]{coverdirection.png}
\end{figure}

\subsection{Geodesics on $\bUs$}
The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ are identical in the sense that their images are in a one-to-one correspondence. Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$. A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$. A geodesic has an initial trajectory defined at 0, represented by a unit vector $v_0\in\RR^2$. If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cO\cup\cE$, the geodesic has a \emph{rational} initial trajectory. For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:

@@

These theorems essentially tell us that these kinds of geodesics behave identically on the Necker cube surface and its cover. Now we show that this is true for $\btUs$ and $\btU$. For these proofs we will refer to the induced homomorphisms from ($\ref{eq:commute}$):

@@

>>>>

\title{Periodicity of Geodesics on the Necker Cube Surface}
\author{Pavel Javornik}

@@

\begin{abstract}
\noindent Although it is most recognizable as M.C. Escher's adaptation of an optical illusion popularized by Louis Albert Necker, the Necker cube surface has a complex structure and is, in analaytic language, described as an infinite, flat (locally isometric to $\RR^2$ on smooth sections), $\RR^3$-embeddable Euclidean cone surface with countably many singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$. Its geometric construction is described as the gluing of infinitely many unit cubes along their edges so that every face shares an edge with exactly four others. A unit-speed geodesic on this surface has an initial trajectory angle $\theta\in\RR/2\pi\ZZ$ given up by the surface's local plane isometry. Given a rational vector direction in $\RR^2$ (components are relatively prime integers) obtained from $\theta$, one can classify every periodic/drift-periodic rational geodesic on the surface. In addition its total period/length and, in the case where it is drift-periodic, translational drift as Euclidean distance in its ambient space is given by solving a system of simple linear Diophantine equations. 
\end{abstract}
\begin{figure}[H]
\centering
\includesvg[width=4in]{cubesxyzpaths2bwv2b}
\label{fig:front}
\caption{Periodic and drift-periodic flows on the Necker cube surface}
\end{figure}

@@

The Necker cube has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems before and after appearing in the works of M.C. Escher [cite]. Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometric properties [cite]. The solid presentation of the cube, when rendered as a flat surface with three rear faces removed, achieves a similar effect when its three visible faces are shaded in a particular way. The Necker cube surface is then obtained by gluing infinitely many copies along its outer edges as in Figure 1.

Dynamical classification of geodesics on flat, periodic surfaces such as the Necker cube surface ultimately boils down to  studying the surface's various symmetries and how the rotational holonomy group acts on the unit tangent bundle of the surface as the unit vector (derivative of the geodesic) is parallel transported along its path. Loops on polyhedra punctured at its cone singularities tend to have discrete holonomy groups as its relative holonomy (holonomy of contractible loops) is trivial.

@@

\subsection{Discussion of Results}
Our initial experiments strongly supported the theory that there would be a correlation between a choice of trajectory angle and dynamical properties of a geodesic on such a symmetric object. Rightfully so, a surface composed of infinitely many cubes an induced flat metric where every neighborhood is locally isometric to the Euclidean plane. Via the parallel transport of a unit tangent vector over the sharp edges of the surface denoted $\bUs$, it becomes necessary to think of the geodesic as a sequence of line segments contained in the faces of each cube on the surface's embedded form in $\RR^3$. $\ref{fig:front}$.
\\
\begin{wrapfigure}{l}{2in.}
\includesvg{vectorplane}
\end{wrapfigure}
\noindent Let $u_0$ be a point on the surface that is contained in the interior of a face and consider a tangent unit vector $v\in\RR^3$ based at point $u_0$. Each face is parallel to a $2$-dimensional subspace of $\RR^3$ spanned by exactly two basis vectors, so one component of $v$ is 0. Call this projected vector $v_0\in\RR^2$.
\\\\\\
The angle $v_0$ makes relative to a choice of basis is the \emph{initial trajectory angle} of the unit-speed geodesic on the surface, $\Phi_t:\RR\rightarrow\bUs$. A rational initial trajectory angle in $\RR^2$ falls into one of two categories:
\begin{Def}
Let $v_0$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ with $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$. We say $v_0$ is an \textbf{odd-odd} vector if its components are relatively prime and both odd. We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$. We say that $v_0$ is an \textbf{even-odd} vector if its components are relatively prime and of opposite parity. We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
\end{Def}
\noindent For convenience we call the disjoint union of these sets $\cS$. This paper aims to depict the relationship between initial trajectory angles and dynamical properties of a geodesic on the surface. Denote $\bUs$'s translational isometries by $\text{Trans}(\bUs)$. Recall that a geodesic on the surface is \emph{periodic} if there exists some $T>0$ such that $\Phi(t+T)=\Phi(t)$ for all $t\in\RR$ and a geodesic is \emph{drift-periodic} if there exists some $T'>0$ and non-trivial $p\in \text{Trans}(\bUs)$ such that $\Phi(t+T')=p(\Phi(t))$ for all $t\in\RR$. In either case we call this value the \emph{period} of $\Phi$ if it is the \emph{smallest} possible value that satisfies our definition. We additionally require that at no point does $\Phi$ encounter a cone singularity of the surface, in which case the geodesic behavior is non-deterministic.

\begin{align}
\text{Let }\XX'=\left\< \left[ \hspace{1mm} \begin{matrix}
							1 & 2\\
							 0 & 1
						\end{matrix}\hspace{1mm}\right],~
\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							 2 & 1
						\end{matrix}\hspace{1mm}\right]\right\>
\end{align}

\noindent Section [blank] explains the significance of this group. Let $M=\left[ \hspace{1mm} \begin{matrix}
							s & u\\
							 w & z
						\end{matrix}\hspace{1mm}\right]\in\XX'$.

\begin{thm*}
(Dynamical Properties of a Geodesic on $\bUs$) \\Let $\Phi_t:\RR\rightarrow\bUs$ be a non-singular unit-speed geodesic on the Necker cube surface such that $\Phi(0)=u_0$ and $\Phi'(0)=v\in\RR^3$, isometric to $v_0\in\RR^2$. Let $\overline{u}_t\in\RR^3$ be $\Phi_t$'s coordinate in $\bUs$'s ambient space at time $t$. Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item $\Phi$ is periodic if $v_0\in\cO$.
\item Suppose $(i)$ is true. If $M (kv_0)=(1,1)$ then $\Phi$ has period $T=6\sqrt{2}||kv_0||$.
\item $\Phi$ is drift-periodic if $v_0\in\cE$. 
\item Suppose $(iii)$ is true. If $M (kv_0)=(1,0)$ or $M (kv_0)=(0,1)$ then $\Phi$ has period $T=2||kv_0||$, and $d_E(\overline{u}_t,\overline{u}_{t+T})=||kv_0||\sqrt{2}$.
\end{enumerate}
\end{thm*}
\compav{Remember to prove that such an M exists.}

Here $||~||$ is the standard Euclidean norm in $\RR^2$, and $d_E:\RR^3\times\RR^3\rightarrow\RR$ is the Euclidean distance. The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.

@@

\section{Periodic Tiling of Necker Cubes}
This section will detail how the Necker Cube surface is constructed as a covering space of a thrice-punctured torus. Make the following identifications on 3 unit squares:
\begin{figure}[H]
\centering
\includesvg[width=1.83in]{neckercube}\includesvg[width=2in.]{neckercubepath}
\caption{$\bGs$ with and without paths $a,b,c,d$.}
\end{figure}

We will call this ``half-cube" $\bG$. It is a genus one piecewise smooth surface homeomorphic to $\TT^2$. Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$. Let $\Sigma_{\bG}\subset\bG$ be the set of these three vertices. These are conical singularities of the surface. We denote the surface punctured at these points by $\bG\backslash\Sigma_{\bG} = \bGs$. The paths labeled $a,b,c,d$ are independent of one another and span $\bGs$'s first fundamental group, $\pi_1(\bGs)\cong \<a,b,c,d \>$, the free group of four generators. The unpunctured Necker cube surface $\bU$ is embedded in $\RR^3$ by tiling along $\bG$'s outer edges:

\begin{figure}[H]
\centering
\includesvg[width=4in]{neckercubepathsurface}
\end{figure}
From this diagram we see that $\bU$ is a universal cover of $\TT^2$. In this paper we are studying geodesic behavior on $\bU$ and in order to do so, it is necessary to determine how the parallel transport of a unit vector around a cone singularity acts on the unit tangent bundle of the surface. But to study the path monodromy we require that these loops (see $c,d$ in Fig 3) aren't contractible on the cube's sharp corners. Denote the countably infinite set of singularities of $\bU$ by $\Sigma_{\bU}$. The \emph{Necker cube surface} is the punctured surface $\bUs=\bU\backslash\Sigma_{\bU}$. Observe from the previous diagram that $\pi_1(\bUs)$ is the kernel of the following group homomorphism:

\begin{Def}
Denote $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism such that $c,d\mapsto(0,0)$, $a\mapsto(1,0)$, and $b\mapsto(0,1)$.
\end{Def}

Take $\pi_1(\bUs)$ to be $\ker\varphi_1$. Then $\Delta_{\varphi_1}=\pi_1(\bGs)/\pi_1(\bUs)\cong\ZZ^2$ is the \emph{deck group} of the covering map $\varphi_1^*:\bUs\rightarrow\bGs$. Now $\bGs$ is unfolded and flattened in order to study monodromy as it transports a vector in $\RR^2$ along a path. Apply the following cutting and gluing operations to the surface: \\
\begin{figure}[H]
\centering
\includesvg[width=3in]{LTorus}
\end{figure}
\noindent This L-shaped torus is then taken apart and recovered as $2\times 2$ torus missing a unit square with the following identifications:\\
\begin{figure}[H]
\centering
\includesvg[width=2in]{BranchTorus}
\caption{$\bGs$ and images of paths after reconstruction. The outermost vertex is \emph{not} a puncture.}
\end{figure}

Consider a $2\times2$ sheet missing an open unit square and that square's vertices (punctures). Center this structure around $0\in\CC$ and recover $\bGs$ with the identifications made in Fig 3. Let $m,n\in\ZZ$. Similar identifications can be made on the following subset of $\CC$ to recover $\bUs$:

\begin{equation}
\label{eq:P}
\mathbf{P} = \mathbb{C}\text{ }\backslash\bigcup_{m,n \in {\mathbb Z}} \big{\{} u+vi:~|u-2m|<\frac{1}{2},~|v-2n|<\frac{1}{2},~u=2m\pm\frac{1}{2}~v=2n\pm\frac{1}{2} \big{\}}.
\end{equation}

\noindent Let $\aleph_{m,n}=2m+i2n\in\CC$ be the center of a removed square. Now let $\zeta_{m,n}(t)=(2m+t)+i(2n-\frac{1}{2})$ and 
$\xi_{m,n}(t)=(2m+t)+i(2n+\frac{1}{2})$ be paths on $\mathbf{P}$ that run along the edges of the squares, parameterized by $t\in (-\frac{1}{2},\frac{1}{2})$ and $(m,n)\in\ZZ^2$. $\bUs$ is recovered as a topological quotient $\mathbf{P}/\sim$, where $\sim$ is the minimal relation on $\mathbf{P}$:

\begin{equation}
\begin{split}
\zeta(t)\sim e^{i\frac{\pi}{2}}~\overline{\zeta(t)-\aleph}\\
\eta(t)\sim e^{i\frac{\pi}{2}}~\overline{\eta(t)-\aleph}
\label{eq:rel2}
\end{split}
\end{equation}
The translations, $\pm 2,\pm 2i$, on $\mathbf{P}$ are isometries that preserve these identifications and induce isometries of $\bUs$. The group generated by these horizontal and vertical translations form the \emph{group of Deck transformations}, $\text{Deck}(\varphi_1^*)$, that  is isomorphic to $\ZZ^2$:

@@

Identify 2 with $f_h$, and $2i$ with $f_v$. Let $x_0\in\bGs$ and fix a point in the fiber,  $u_0\in\varphi_1^*{-1}(x_0)\subset\bUs$. The action of $\ZZ^2$ on the fiber over $x_0$ is given by $(m,n)\cdot u_0=(f_h^m\circ f_v^n)(u_0)$. The distance between $u_0$ and $(f_h^m\circ f_v^n)(u_0)$ is measured in its ambient space and taken to be the magnitude of $2m+i2n\in\CC$.

\begin{lem}
Let $[\alpha]$ be homotopy class not in the kernel of $\varphi_1$. It will lift to an unclosed path terminating on $\varphi_1([\alpha])\cdot u_0$.
\end{lem}

\subsection{Monodromy Group Representation}
A \emph{rotational monodromy map} is a morphism from the first fundamental group of a surface to the \emph{monodromy group} acting on the unit tangent bundle via parallel transport of a vector along its path. Existence of a parallel transport map is a consequence of the metric connection on a surface with complex structure. The monodromy map is well defined when the \emph{holonomy group} is \emph{discrete}. When the \emph{relative holonomy} of contractible loops is trivial, the holonomy group is the set of attainable rotations of a unit vector upon returning to the base point after transporting it along a non-trivial loop. A \emph{flat surface} like $\bGs$ has trivial relative holonomy as its neighborhoods are isometric to the plane. Denote the \emph{unit tangent bundles} of $\bUs$ and $\bGs$ by $T^1\bUs$ and $T^1\bGs$. If $\alpha$ is a loop on $\bGs$, then $[\alpha]$ acts on a fixed vector $v_0\in\RR^2$ by rotation. In the case where $\alpha$ is \emph{not} a geodesic segment on the surface of length $L$, $v_0=\alpha'(0)\in\RR^2$ may differ from $\alpha'(L)$ by some rotation.

\begin{Def}
Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ to be the group homomorphism where $a,b \mapsto I_2$, $c  \mapsto R$, and $d  \mapsto R^3$, where $I_2$ is the identity matrix and $R$ is a positive $90^\circ$ rotation matrix.
\end{Def}

\begin{figure}[H]
\centering
\includegraphics[width=2.5in]{monodromy.png}
\caption{Effect a nontrivial loop has on basis vectors $(i,j)\in\RR^2$.}
\label{fig:loop}
\end{figure}

\begin{lem}
$\varphi_2:\pi_1(\bGs)\rightarrow \mathrm{SO}(2,\ZZ)$ is the rotational monodromy representation of $\bGs$.
\begin{proof}
From Figure 4 it's clear that $c,d$ will rotate a unit vector by $\pm 90^\circ$ by parallel transport along their paths. These values are determined by the cone angles of the singularities they loop around. The parallel transport of a vector along $a$ or $b$ would be trivial since the outer edges they pass through are glued by translation.
\end{proof}
\end{lem}

\begin{lem}
Any element in $\text{Deck}(\varphi_1^*)$ acting on $T^1\bUs$ has a trivial effect on holonomy.
\begin{proof}
Since $\bGs=\bUs/\ZZ^2$, the deck group action lifts naturally to an action on their tangent vector spaces and unit tangent bundles via pushforward of deck transformations $f_h,f_v$ on the bundle maps. i.e. $T^1\bGs=T^1\bUs/\ZZ^2$. This effect must be trivial since $\varphi_1(a)\sim f_h$ and $\varphi_1(b)\sim f_v$, and both homotopy classes $a,b$ have trivial holonomy.  \\
How do $f_h,f_v$ parallel transport a vector?\\
Well if $\varphi_1([a])\cdot u_0=f_h(u_0)$, then for 
\end{proof}
\end{lem}

\begin{thm}
Let $m=\varphi_2\vert_{\pi_1(\bUs)}$. Then $m$ is the rotational monodromy map of $\bUs$.
\begin{proof}

\end{proof}
\end{thm}

\begin{cor}
Let $\alpha:[0,1]\rightarrow\bGs$ be a loop on $\bGs$ with trivial holonomy and $\tilde{\alpha}$ its lift to $\bUs$. Then $\tilde{\alpha}$ has trivial holonomy as well.
\begin{proof}

\end{proof}
\end{cor}

@@

\subsection{A Four-Fold Cover}
We now take these covers we built to construct a third one that covers them all. This is a four-fold cover of $\bUs$ and a $\ZZ^2$ translation cover of $\bMs$. 
\begin{Def}
Let $\btUs$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUs)=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$.
\end{Def}

It follows that $\ker\varphi_1\big\vert_{\pi_1(\bMs)}=\ker\varphi_2\big\vert_{\pi_1(\bUs)}$ and that $\btUs$ is a cover of these two surfaces. Edge identifications are made on four copies of $\mathbf{P}$, possibly indexed by the finite component of the direct product $\mathbf{P}\times\ZZ/4\ZZ$. The Deck group of the covering map from $\btUs$ to $\bGs$ is the set $\ZZ^2\times \SO(2,\ZZ)$. Consider a loop on $\bUs$:

@@

\noindent If we lift this curve to $\btUs$ we can identify it with an element of the deck group, and check if its lift from $\bGs$ belongs to the kernel of both $\varphi_1$ and $\varphi_2$. Fix a direction on this surface pointing north that respects $\bUs$'s compass and rotate every plane by either $90^\circ,$ $180^\circ,$ or $270^\circ$ as the curve travels from one plane (copy of $\mathbf{P}$) to the next. Observe that it will not only close if the curve acts trivially on the holonomy of $\bUs$, but with this construction we can see that $\btUs$ is actually an infinite-type translation surface:

\begin{figure}[H]
\centering
\includegraphics[width=3in]{coverdirection.png}
\caption{With $P\in 2\ZZ+2\ZZ i$.}
\end{figure}

\noindent

\subsection{Geodesics on $\bUs$}
The key connection made between surfaces $\bUs$ and $\btUs$ is that geodesics on $\bUs$ and $\btUs$ behave identically. Unless stated otherwise, a \emph{geodesic} is a \emph{unit-speed} locally distance minimizing curve on a smooth topological surface as a function parameterized by an element in $\RR$. A \emph{geodesic segment} is a restriction of the geodesic to some closed interval in $\RR$. A geodesic has an initial trajectory defined at time 0, represented by a unit vector $v_0\in\RR^2$. If the vector can be scaled by $k\in\RR$ such that $kv_0\in\cS$, the geodesic has a \emph{rational} initial trajectory. For example, the following is a periodic geodesic on the surface with initial direction $kv_0\in\cO$:

@@

These theorems essentially tell us that these types of geodesics behave identically on the Necker cube surface and its cover. Now we show that this is true for $\btUs$ and its metric completion, $\btU$.

@@

This follows readily from the fact that geodesics are undefined on cone singularities. A geodesic on $\btUs$ can be mapped bijectively to a geodesic on  $\btU$ and vice-a-versa.
